Going from coarse landings data to fine scale species distribution

### 2022-03-30 Version_M2_plusieurszones/draft/paper_3/paper_3.Rmd
paper_3

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

@@

```{r setup, include = FALSE}
library(cowplot)
library(doBy)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(kableExtra)
library(latex2exp)
library(mapdata)
library(scales)
library(papaja)
library(rnaturalearth)
library(sf)
```

@@

All fishing locations $x_i$ are supposed to be known through VMS data. At each fishing location $x_i$, fishermen realize a catch $Y_i$ conditionnally on the latent field values (referred hereafter as a punctual catch). Observations are supposed to follow some distribution $\mathcal{L}_Y$. In our case, observations are zero-inflated positive continuous data and they are modelled through a zero-inflated lognormal model previously introduced by @thorson_three_2018 and already used in @alglave_combining_2022.

@@

$\mathcal{L}_Y$ is decomposed it 2 parts:

@@

- if the catch is positive, the probability to obtain some observation value $y_i$ is modelled through a continous distribution $\operatorname{L}$ (here a lognormal distribution - its parameterization is described in SM...)) with mean component $\frac{S(x_{i})}{1-p_{i}}$ and variance term $\sigma^2$.

@@

Fishermen must declare what they catch at the level of the statistical rectangle. Declarations are daily recorded in logbooks data. The declaration (denoted $D_j$) is a summation of all the $Y_i$ realized on the fishing positions belonging to the catch declaration $D_j$ and they can be expressed as:

@@

$\mathcal{P}_j$ is the vector of the fishing observations belonging to declaration $D_j$. The punctual observations $Y_i$ are indexed through $i \in [\![1,m_j]\!]$ with $m_j$ the number of fishing positions belonging to the $j^{th}$ declaration.

@@

In standard processing, $D_j$ (from logbooks) are reallocated uniformly on related $x_i$ (available from VMS) so that derived punctual observations $Y^*_i$ are computed as $Y^*_i=D_j/m_j$. This is what we call uniform (or proportional) reallocation. In this case, inference of species distribution is directly computed through reallocated $Y^*_i$ assuming these are the exact punctual observations. This strongly simplifies the actual process of observation and have most likely repercussion on model performance.

@@

with $S_{\mathcal{P}_j}=(S(x_1),...,S(x_i), ...,S(x_{m_j}))$, $\xi$ the zero-inflation parmeter and $\sigma^2$ the variance parameter. As $D_j=\sum_{i \in \mathcal{P}_j}{Y_{i}}$, it is possible to relate $D_j$, $Y_i$ and $S_{\mathcal{P}_j}$ through the distribution $\mathcal{L}_Y$ and $\mathcal{L}_D$. Our approach is to match the moments of $D_j$ (obtained from the moments of $\mathcal{L}_Y$) and $\mathcal{L}_D$ and then to infer $S(x)$ and $Y_i$ from $D_j$. This is what we call model-based reallocation. This model will be referred as the $D_j$ model. In contrast, the model fitted on punctual observations $Y_i$ will be referred as the $Y^*_i$ model.

In the following, $Y$ and $D$ will be assumed conditionnal on the fishing positions and the latent field values, but for notation simplicity they will be simply denoted as $Y_i$ or $D_j$ instead of $Y_i \vert x_i, S(x_i)$ and $D_j | S_{\mathcal{P}_j},\mathcal{P}_j$.

@@

If we consider $\mathcal{L}_{D_j|D_j>0}$ is lognormal (with parameterization described in SM...) then, $\mu_j = E(D|D_j>0)$ and $\sigma_j = ln(\frac{Var(D|D_j>0)}{E(D|D_j>0)^2} + 1)$. All calculations to obtain these formulas are available in SM. By fitting the model on declaration $D_j$ rather than reallocated catch $Y_i$ and linking $\mathcal{L}_D$ to $\mathcal{L}_Y$, we intend to better follow the observation process of the commercial data and potentially improve inference compared with the approximation realized through uniform reallocation.

@@

![(\#fig:SchemeRealloc) Reallocation process.](images/realloc.png)

@@

To evaluate what bring the performance of the new approach, we assess models performance through simulation-estimation.

First, we conduct basic simulations in order to explore the base properties of the framework and the effect of some key variables on model performance. We assume the study domain is a single statistical rectangle, only commercial data feed the model and the expression of the latent field in both simulation and estimation is simplified (the variability only depends on one covariate, not on a spatial random term). These simulations will be referred as single-square simulations.

@@

In these two sets of simulation-estimation, the covariate is modelled as a continous GRF that we suppose known at each point of the grid. The covariate effect is fixed to $\beta_S=2$ and the intercept is also fixed to $\mu=2$. Regarding commercial data, the number of fishing pings per declaration is fixed to 10 as it is the average number of fishing locations for a single declaration in real data. All parameterizations are detailed in the table ....

@@

@@

- a golden standard configuration: observations are known exactly and the $Y^*_i$ model is fitted to the exact data.

- a configuration corresponding to the actual situation (uniform reallocation): simulated catches are summed into declarations and reallocated over the related fishing locations. We then fit the $Y^*_i$ model to the reallocated data ($Y_i^*$).

- a configuration corresponding to our alternative approach: simulated catch are reallocated over fishing locations and we fit the $D_j$ model to the data (i.e. the one working at the level of $D_j$).

@@

The study area is based on the case study; it includes the whole coast of the Bay of Biscay (Figure \@ref(fig:MapSeveral)A). To tailor the case study, we simulate 3000 of fishing positions regrouped in 300 declarations (10 pings per declaration). Commercial data may not cover the full area, and consequently we allow the commercial samples to cover only 2/3 ot the area. Similarly to the single-square simulations, the sampling of the commercial points of a declaration are realized in 3 steps. (1) The declaration is randomly affected to one of the ICES rectangle. (2) The centroid of the fishing zone is randomly sampled within this statistical rectangle. (3) The 10 fishing locations are randomly sampled within the fishing zone. The radius of the fishing zone is set so as the extent of a fishing operation does not exceed 30 km. Note that we do not explore the effect of exploring several zones within the same declaration as it is already done in the single-square simulations.

100 scientific samples are simulated following a random stratified plan; contrarly to commercial data they cover the entire study domain (Figure \@ref(fig:MapSeveral)A). Scientific observations are simulated following the observation equations of $\mathcal{L}_Y$ (with specific parameters for scientific data - Table ...).

@@

- to assess what bring our alternative approach, we compare the $Y^*_i$ model and the $D_j$ model.

@@

@@

The $D_j$ model faced congvergence issues (some of the parameters were hardly estimated e.g. the range parameter). To ease convergence, we integrated in the analysis onboard observer data for the same fleet. They are considered as exact commercial data of catches. Then, integrating these data allow to have direct information on $Y_i$ and to better estimate the observation equations parameters (i.e. observation variance and zero-inflation parameter of commercial data).

@@

## Single square analysis

@@

## Table of convergence
Results_conv <- Results
Results_conv$one <- 1
Results_conv_2 <- summaryBy(Convergence+one~
                              sequencesdepeche+
                              # zonespersequence+
                              aggreg_obs+
                              # b_true+
                              n_samp_com+
                              reallocation,
                            data=Results_conv,
                            FUN=sum) %>%
  dplyr::select(n_samp_com,
                sequencesdepeche,
                # zonespersequence,
                reallocation,
                aggreg_obs,
                Convergence.sum,
                one.sum) %>%
    mutate(aggreg_obs = ifelse(aggreg_obs == T,"Dj","Yi"),
           reallocation = ifelse(reallocation == 1,"Yes","No"),
           perc_convergence = round((1 - Convergence.sum / one.sum)*100,digits = 3))

@@

knitr::kable(Results_conv_2,booktabs = T,align = "c",
                  caption = "Percentage of convergence per simulation/model configuration at the level of a single rectangle")

@@

Results <- Results %>%
  mutate(relloc_aggreg = paste0("Reallocation: ",ifelse(reallocation==0,"No","Yes")," -  Likelihood: ",ifelse(aggreg_obs==T,"Dj","Yi")))

Results$zonespersequence <- as.factor(Results$zonespersequence)
Results$b_true <- as.factor(Results$b_true)
Results$relloc_aggreg <- factor(Results$relloc_aggreg,
                                levels = c("Reallocation: No -  Likelihood: Yi",
                                           "Reallocation: Yes -  Likelihood: Yi",
                                           "Reallocation: Yes -  Likelihood: Dj"))

@@

MSPE_S_plot <- ggplot()+
  geom_boxplot(data = Results_2,
               aes(x = zonespersequence,
                   y = mspe,
                   fill = relloc_aggreg),
               outlier.colour = NA)+
  theme_bw()+
  facet_wrap(.~factor(n_samp_com))+
  theme(legend.title = element_blank(),
        legend.position = "none",
        aspect.ratio = 1)+
  ylab("MSPE")+
  xlab("")+
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
                labels = trans_format("log10", math_format(10^.x)))+
  scale_fill_manual(breaks = c("Reallocation: No -  Likelihood: Yi",
                                "Reallocation: Yes -  Likelihood: Yi",
                                "Reallocation: Yes -  Likelihood: Dj"),
                     values=c("gold","#F8766D","#00BA38"))

Beta_plot <- ggplot()+
  geom_boxplot(data = Results_2,
               aes(x = zonespersequence,
                   y = beta1_est,
                   fill = relloc_aggreg),
               outlier.colour = NA)+
  geom_hline(yintercept = 0,linetype="dashed")+
  geom_hline(yintercept = 2,col="red",linetype="dashed")+
  theme_bw()+
  facet_wrap(.~factor(n_samp_com))+
  theme(legend.title = element_blank(),
        legend.position = "none",
        aspect.ratio = 1)+
  # ylim(-1,1)+ 
  ylab(parse(text = TeX('$\\beta_S$')))+
  xlab("")+
  scale_fill_manual(breaks = c("Reallocation: No -  Likelihood: Yi",
                                "Reallocation: Yes -  Likelihood: Yi",
                                "Reallocation: Yes -  Likelihood: Dj"),
                     values=c("gold","#F8766D","#00BA38"))

RelBias_N_plot <- ggplot()+
  geom_boxplot(data = Results_2,
               aes(x = zonespersequence,
                   y = RelBias_N,
                   fill = relloc_aggreg),
               outlier.colour = NA)+
  geom_hline(yintercept = 0,linetype="dashed")+
  theme_bw()+
  facet_wrap(.~factor(n_samp_com))+
  theme(legend.title = element_blank(),
        aspect.ratio = 1,
        legend.position = 'none')+
  ylim(-1,1)+
  ylab("Relative bias of biomass")+
  xlab("")+
  scale_fill_manual(breaks = c("Reallocation: No -  Likelihood: Yi",
                                "Reallocation: Yes -  Likelihood: Yi",
                                "Reallocation: Yes -  Likelihood: Dj"),
                     values=c("gold","#F8766D","#00BA38"))

@@

q1_plot <- ggplot()+
  geom_boxplot(data = Results_3,
               aes(x = relloc_aggreg,
                   y = q1_com_est,
                   fill = relloc_aggreg),
               outlier.colour = NA)+
  geom_hline(yintercept = 0,linetype="dashed")+
  geom_hline(yintercept = -1,col="red",linetype="dashed")+
  theme_bw()+
  theme(legend.title = element_blank(),
        legend.position = "none",
        aspect.ratio = 1,
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(breaks = c("Reallocation: No -  Likelihood: Yi",
                               "Reallocation: Yes -  Likelihood: Yi",
                               "Reallocation: Yes -  Likelihood: Dj"),
                    values=c("gold","#F8766D","#00BA38"))+
  xlab("")+ylab(parse(text = TeX('$\\xi$')))+
  ggtitle("Zero-inflation parameter")

sigma_plot <- ggplot()+
  geom_boxplot(data = Results_3,
               aes(x = relloc_aggreg,
                   y = sigma_com_est,
                   fill = relloc_aggreg),
               outlier.colour = NA)+
  geom_hline(yintercept = 0,linetype="dashed")+
  theme_bw()+
  theme(legend.title = element_blank(),
        legend.position = "none",
        aspect.ratio = 1,
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(breaks = c("Reallocation: No -  Likelihood: Yi",
                               "Reallocation: Yes -  Likelihood: Yi",
                               "Reallocation: Yes -  Likelihood: Dj"),
                    values=c("gold","#F8766D","#00BA38"))+
  xlab("")+ylab(parse(text = TeX('$\\sigma^2$')))+
  ggtitle("Observation variance parameter")+
  geom_hline(yintercept = 1,col="red",linetype="dashed")

intercept_plot <- ggplot()+
  geom_boxplot(data = Results_3,
               aes(x = relloc_aggreg,
                   y = intercept_est,
                   fill = relloc_aggreg),
               outlier.colour = NA)+
  geom_hline(yintercept = 0,linetype="dashed")+
  theme_bw()+
  theme(legend.title = element_blank(),
        legend.position = "none",
        aspect.ratio = 1,
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(breaks = c("Reallocation: No -  Likelihood: Yi",
                               "Reallocation: Yes -  Likelihood: Yi",
                               "Reallocation: Yes -  Likelihood: Dj"),
                    values=c("gold","#F8766D","#00BA38"))+
  xlab("")+ylab(parse(text = TeX('$\\mu$')))+
  ggtitle("Intercept of the latent field")+
  geom_hline(yintercept = 2,col="red",linetype="dashed")

@@

![(\#fig:PerfMetricSingle) Performance metric for single-square simulations. columns: number of commercial pings. x-axis: number of zones visited within each declaration. 'Reallocation: ', data are or are not reallocated in simulations. 'Likelihood: ', the likelihood is computed on reallocated observations $Y_i$ or on catch declarations $D_j$. Gold: golden standard. Gold: golden standard. Red: uniform reallocation situation. Green: model-based reallocation ($D_j$ model). Simulations conducted on 10 fishing positions are not represented as they mostly did not converged for the $D_j$ model.](images/Perf.metric_single_square.png)

@@

The $D_j$ model allows to recover the species-habitat relationship and to improve the accuracy of the spatial predictions (Figure \@ref(fig:PerfMetricSingle)). Still, the $D_j$ model outputs are not as accurate as the ones of the golden-standard. Furthermore, the zero-inflation parameter is unbiased when the model is fitted to catch declarations $D_j$. Other parameters (observation variance, intercept) are also better estimated even though they remain slightly biased (Figure \@ref(fig:ParBiasSingle)). This alternative model have some convergence difficulties (Table 1) as 8% of the model did not converged when sample size is medium (100 pings) and only 3% did not when sample size is large (1000 pings).

![(\#fig:ParBiasSingle) Parameters relative bias for single-square simulations. 'Reallocation: ', data are or are not reallocated in simulations. 'Likelihood: ', the likelihood is computed on reallocated observations $Y_i$ or on catch declarations $D_j$. Gold: golden standard. Red: uniform reallocation situation. Green: model-based reallocation ($D_j$ model). Only the simulations with 1000 fishing positionss are represented. Black line: zero value. Red line: parameter true value.](images/par_plot_single_square.png)

## Multiple square analysis

@@

@@

@@

Results_conv_2 <- Results_conv_2[-which(Results_conv_2$Model == "Scientific model" &
                                          Results_conv_2$`Likelihood level` == "Dj"),]
Results_conv_2$`Likelihood level`[which(Results_conv_2$Model == "Scientific model" &
                       Results_conv_2$`Likelihood level` == "Yi")] <- " "
knitr::kable(Results_conv_2,booktabs = T,align = "c",
                  caption = "Percentage of convergence per simulation/model configuration at the level of several ICES rectangles")

@@

Results_plot$obs <- NA
Results_plot$obs[which(Results_plot$aggreg_obs == T)] <- "Dj"
Results_plot$obs[which(Results_plot$aggreg_obs == F)] <- "Yi"
Results_plot$obs[which(Results_plot$Estimation_model == 2)] <- "Scientific data"
Results_plot$obs <- factor(Results_plot$obs,levels = c("Yi","Dj","Scientific data"))

@@

beta_plot <- ggplot()+
  geom_boxplot(data = Results_plot,
               aes(x = Model,
                   y = beta1_est,
                   fill = obs))+
  geom_hline(yintercept = 0,linetype="dashed")+
  theme_bw()+
  theme(legend.title = element_blank(),
        legend.position = "none",
        aspect.ratio = 1)+
  # facet_wrap(.~simu_type)+
  ylab(TeX("$\\beta_S$")) +
  xlab("")+
  geom_hline(yintercept = 2,col="red",linetype="dashed")+
  scale_color_manual(breaks = c("Scientific model",
                                "Yi",
                                "Dj"),
                     values=c("#619CFF","#F8766D","#00BA38"))

>>>>

floatsintext     : yes
figurelist       : no
tablelist        : no
footnotelist     : no
linenumbers      : yes
mask             : no
draft            : no

@@

```{r setup, include = FALSE}
library(cowplot)
library(doBy)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(kableExtra)
library(knitr)
library(latex2exp)
library(mapdata)
library(scales)
library(papaja)
library(rnaturalearth)
library(sf)
```

@@

All fishing locations $x_i$ are supposed to be known through VMS data. At each fishing location $x_i$, fishermen realize a catch $Y_i$ conditionnally on the latent field values (referred hereafter as a punctual catch). Observations are supposed to follow some distribution $\mathcal{L}_Y$.

@@

In our case, observations are zero-inflated positive continuous data and they are modelled through a zero-inflated lognormal model previously introduced by @thorson_three_2018 and already used in @alglave_combining_2022. $\mathcal{L}_Y$ is decomposed it 2 parts:

@@

- if the catch is positive, the probability to obtain some observation value $y_i$ is modelled through a positive continous distribution $\operatorname{L}$ (here a lognormal distribution - its parameterization is described in SM)) with mean component $\frac{S(x_{i})}{1-p_{i}}$ and variance term $\sigma^2$.

@@

Fishermen must declare what they catch at the level of the statistical rectangle at a daily time step. These data are recorded in logbooks data. A declaration (denoted $D_j$) is a summation of all the $Y_i$ realized on the fishing positions belonging to the catch declaration $D_j$. They can be expressed as:

@@

$\mathcal{P}_j$ is the vector of the fishing observations belonging to the declaration $D_j$. The punctual observations $Y_i$ are indexed through $i \in [\![1,m_j]\!]$ with $m_j$ the number of fishing positions belonging to the $j^{th}$ declaration.

@@

In standard processing, $D_j$ (available from logbooks) are reallocated uniformly on related $x_i$ (available from VMS) so that derived punctual observations $Y^*_i$ are computed as $Y^*_i=D_j/m_j$. This is what we call uniform (or proportional) reallocation. In this case, inference of species distribution is directly computed through reallocated $Y^*_i$ assuming these are the exact punctual observations. This strongly simplifies the actual process of observation and have most likely repercussion on model performance.

@@

with $S_{\mathcal{P}_j}=(S(x_1),...,S(x_i), ...,S(x_{m_j}))$, $\xi$ the zero-inflation parmeter and $\sigma^2$ the variance parameter. As $D_j=\sum_{i \in \mathcal{P}_j}{Y_{i}}$, it is possible to relate $D_j$, $Y_i$ and $S_{\mathcal{P}_j}$ through the distribution $\mathcal{L}_Y$ and $\mathcal{L}_D$. Our approach is to match the moments of $D_j$ (obtained from the moments of $\mathcal{L}_Y$) and $\mathcal{L}_D$ and then to infer $S(x)$ and $Y_i$ from $D_j$. This is what we call model-based reallocation. This model will be referred as the $D_j$ model. In contrast, the model fitted on reallocated observations $Y^*_i$ will be referred as the $Y^*_i$ model.

In the following, $Y$ and $D$ will be assumed conditionnal on the fishing positions and the latent field values, but for notation simplicity they will be simply denoted as $Y_i$ or $D_j$ instead of $Y_i \vert x_i, S(x_i)$ and $D_j | \mathcal{P}_j,S_{\mathcal{P}_j}$.

@@

If we consider $\mathcal{L}_{D_j|D_j>0}$ is lognormal (with parameterization described in SM) then, $\mu_j = E(D|D_j>0)$ and $\sigma_j = ln(\frac{Var(D|D_j>0)}{E(D|D_j>0)^2} + 1)$. Such way, by fitting the model on declaration $D_j$ rather than reallocated catch $Y_i$ and linking $\mathcal{L}_D$ to $\mathcal{L}_Y$, we intend to better follow the observation process of the commercial data and potentially improve inference compared with the approximation realized through uniform reallocation. All calculations to obtain these formulas are available in SM.

@@

![(\#fig:SchemeRealloc) Schematic representation of the reallocation process.](images/realloc.png)

@@

To evaluate the impact of reallocation on model outputs and how our alternative approach can improve model accuracy, we conduct simulation-estimation.

First, we conduct basic simulations in order to explore the effect of some key variables on model performance and also the base properties of the alternative models. We assume the study domain is a single statistical rectangle, only commercial data feed the model and the expression of the latent field in both simulation and estimation is simplified (the variability only depends on one covariate, not on a spatial random term). These simulations will be referred as "single-square simulations".

@@

In these two sets of simulation-estimation, the covariate is modelled as a continous GRF that we suppose known at each point of the grid. The covariate effect is fixed to $\beta_S=2$ and the intercept is also fixed to $\mu=2$. Regarding commercial data, the number of fishing pings per declaration is fixed to 10 as it is the average number of fishing locations for a single declaration in real data. All parameterizations are detailed in the Table 1.

@@

df <- data.frame(v1=c("$\\mu$","$\\beta_S$","$\\text{Range}$","$\\text{Marginal variance}$","$\\xi_{com}$","$\\sigma_{com}$","$k_{com}$","$\\xi_{sci}$","$\\sigma_{sci}$"),
                 v2=c("2","2"," -- "," -- ","-1","1"," -- "," -- "," -- "),
                 v3=c("2","2","1.5","0.5","-1","1","1","0","0.8"))

# italic column headers
colnames(df) <- c("\\text{Parameters}","\\text{Single-square simulations}","\\text{Multiple-square simulations}")

kable(df, "latex", align=c("l","c","c"), booktabs=TRUE, escape = F, caption = 'Parameter values for the simulations')

table_param <- t(data.frame(
  c("$\\mu$","2","2"),
  c("$\\beta_S$","2","2"),
  c("$\\text{Range}$"," -- ","1.5"),
  c("$\\text{Marginal variance}$"," -- ","0.5"),
  c("$\\xi_{com}$","-1","-1"),
  c("$\\sigma_{com}$","1","1"),
  c("$k_{com}$"," -- ","1"),
  c("$\\xi_{sci}$"," -- ","0"),
  c("$\\sigma_{sci}$"," -- ","0.8")
  ))

@@

- a golden standard configuration: punctual observations ($Y_i$) are known exactly and the punctual-data model is fitted to the exact data.

- a configuration corresponding to the actual situation (uniform reallocation): simulated catches are summed into declarations and reallocated over the related fishing locations. We then fit the $Y_i^*$ model (the model fitted to punctual reallocated data).

- a configuration corresponding to our alternative approach: simulated catch are reallocated over fishing locations and we fit the $D_j$ model to the data (i.e. the one fitted at the level of the declarations data).

@@

The study area is based on the case study; it includes the whole coast of the Bay of Biscay and cover several statistical rectangles (Figure \@ref(fig:MapSeveral)A). To tailor the case study, we simulate 3000 of fishing positions regrouped in 300 declarations (10 pings per declaration). Commercial data may not cover the full area, and consequently we allow the commercial samples to cover only 2/3 ot the area. Similarly to the single-square simulations, the sampling of the commercial points of a declaration are realized in 3 steps. (1) The declaration is randomly affected to one of the ICES rectangle. (2) The centroid of the fishing zone is randomly sampled within this statistical rectangle. (3) The 10 fishing punctual observations are randomly sampled within the fishing zone. The radius of the fishing zone is set so as the extent of a fishing operation does not exceed 30 km. Note that we do not explore the effect of exploring several zones within the same declaration as it is already done in the single-square simulations.

100 scientific samples are simulated following a random stratified plan; contrarly to commercial data they cover the entire study domain (Figure \@ref(fig:MapSeveral)A). Scientific observations are simulated following the observation equations of $\mathcal{L}_Y$ (with specific parameters for scientific data - Table 1).

@@

- to assess what brings our alternative approach, we compare the $Y^*_i$ model and the $D_j$ model.

@@

Note that when including 2 datasets in the model, a difference in catchability should be accounted for and one of the two dataset will be considered as reference (see @alglave_combining_2022 for more details on parameterization). Here the reference level is set with scientific data and the parameter $k_{com}$ accounts for the relative ratio between scientific and commercial observation values. In simulations, $k_{com}$ is set to 1 so that both datasources are assumed to have the same catchability.

@@

The $D_j$ model faced convergence issues (some of the parameters were hardly estimated e.g. the range parameter). To ease convergence, we integrated in the analysis onboard observer data for the same fleet. They are considered as exact commercial data of catches. Integrating these data allow to have direct information on $Y_i$ and to better estimate the observation equations parameters (i.e. observation variance and zero-inflation parameter of commercial data).

@@

## Single-square simulations

@@

## Table of convergence
Results_conv <- Results
Results_conv$one <- 1
Results_conv_2 <- summaryBy(Convergence+one~
                              sequencesdepeche+
                              # zonespersequence+
                              aggreg_obs+
                              # b_true+
                              n_samp_com+
                              reallocation,
                            data=Results_conv,
                            FUN=sum) %>%
  dplyr::select(n_samp_com,
                sequencesdepeche,
                # zonespersequence,
                reallocation,
                aggreg_obs,
                Convergence.sum,
                one.sum) %>%
    mutate(aggreg_obs = ifelse(aggreg_obs == T,"Dj","Yi"),
           reallocation = ifelse(reallocation == 1,"Yes","No"),
           perc_convergence = round((1 - Convergence.sum / one.sum)*100,digits = 3)) %>%
  mutate(aggreg_obs = ifelse((aggreg_obs == "Yi" & reallocation == "Yes"),"Yi*",aggreg_obs))

@@

@@

Results <- Results %>%
  mutate(relloc_aggreg = paste0("Reallocation: ",ifelse(reallocation==0,"No","Yes")," -  Likelihood: ",ifelse(aggreg_obs==T,"Dj","Yi"))) %>%
  mutate(relloc_aggreg = ifelse(aggreg_obs == F & reallocation == 1,"Reallocation: Yes -  Likelihood: Yi*",relloc_aggreg))

Results$zonespersequence <- as.factor(Results$zonespersequence)
Results$b_true <- as.factor(Results$b_true)
Results$relloc_aggreg <- factor(Results$relloc_aggreg,
                                levels = c("Reallocation: No -  Likelihood: Yi",
                                           "Reallocation: Yes -  Likelihood: Yi*",
                                           "Reallocation: Yes -  Likelihood: Dj"))

@@

MSPE_S_plot <- ggplot()+
  geom_boxplot(data = Results_2,
               aes(x = zonespersequence,
                   y = mspe,
                   fill = relloc_aggreg),
               outlier.colour = NA)+
  theme_bw()+
  facet_wrap(.~factor(n_samp_com))+
  theme(legend.title = element_blank(),
        legend.position = "none",
        aspect.ratio = 1)+
  ylab("MSPE")+
  xlab("")+
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
                labels = trans_format("log10", math_format(10^.x)))+
  scale_fill_manual(breaks = c("Reallocation: No -  Likelihood: Yi",
                                "Reallocation: Yes -  Likelihood: Yi*",
                                "Reallocation: Yes -  Likelihood: Dj"),
                     values=c("gold","#F8766D","#00BA38"))

Beta_plot <- ggplot()+
  geom_boxplot(data = Results_2,
               aes(x = zonespersequence,
                   y = beta1_est,
                   fill = relloc_aggreg),
               outlier.colour = NA)+
  geom_hline(yintercept = 0,linetype="dashed")+
  geom_hline(yintercept = 2,col="red",linetype="dashed")+
  theme_bw()+
  facet_wrap(.~factor(n_samp_com))+
  theme(legend.title = element_blank(),
        legend.position = "none",
        aspect.ratio = 1)+
  # ylim(-1,1)+ 
  ylab(parse(text = TeX('$\\beta_S$')))+
  xlab("")+
  scale_fill_manual(breaks = c("Reallocation: No -  Likelihood: Yi",
                                "Reallocation: Yes -  Likelihood: Yi*",
                                "Reallocation: Yes -  Likelihood: Dj"),
                     values=c("gold","#F8766D","#00BA38"))

RelBias_N_plot <- ggplot()+
  geom_boxplot(data = Results_2,
               aes(x = zonespersequence,
                   y = RelBias_N,
                   fill = relloc_aggreg),
               outlier.colour = NA)+
  geom_hline(yintercept = 0,linetype="dashed")+
  theme_bw()+
  facet_wrap(.~factor(n_samp_com))+
  theme(legend.title = element_blank(),
        aspect.ratio = 1,
        legend.position = 'none')+
  ylim(-1,1)+
  ylab("Relative bias of biomass")+
  xlab("")+
  scale_fill_manual(breaks = c("Reallocation: No -  Likelihood: Yi",
                                "Reallocation: Yes -  Likelihood: Yi*",
                                "Reallocation: Yes -  Likelihood: Dj"),
                     values=c("gold","#F8766D","#00BA38"))

@@

q1_plot <- ggplot()+
  geom_boxplot(data = Results_3,
               aes(x = relloc_aggreg,
                   y = q1_com_est,
                   fill = relloc_aggreg),
               outlier.colour = NA)+
  geom_hline(yintercept = 0,linetype="dashed")+
  geom_hline(yintercept = -1,col="red",linetype="dashed")+
  theme_bw()+
  theme(legend.title = element_blank(),
        legend.position = "none",
        aspect.ratio = 1,
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(breaks = c("Reallocation: No -  Likelihood: Yi",
                               "Reallocation: Yes -  Likelihood: Yi*",
                               "Reallocation: Yes -  Likelihood: Dj"),
                    values=c("gold","#F8766D","#00BA38"))+
  xlab("")+ylab(parse(text = TeX('$\\xi$')))+
  ggtitle("Zero-inflation parameter")

sigma_plot <- ggplot()+
  geom_boxplot(data = Results_3,
               aes(x = relloc_aggreg,
                   y = sigma_com_est,
                   fill = relloc_aggreg),
               outlier.colour = NA)+
  geom_hline(yintercept = 0,linetype="dashed")+
  theme_bw()+
  theme(legend.title = element_blank(),
        legend.position = "none",
        aspect.ratio = 1,
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(breaks = c("Reallocation: No -  Likelihood: Yi",
                               "Reallocation: Yes -  Likelihood: Yi*",
                               "Reallocation: Yes -  Likelihood: Dj"),
                    values=c("gold","#F8766D","#00BA38"))+
  xlab("")+ylab(parse(text = TeX('$\\sigma^2$')))+
  ggtitle("Observation variance parameter")+
  geom_hline(yintercept = 1,col="red",linetype="dashed")

intercept_plot <- ggplot()+
  geom_boxplot(data = Results_3,
               aes(x = relloc_aggreg,
                   y = intercept_est,
                   fill = relloc_aggreg),
               outlier.colour = NA)+
  geom_hline(yintercept = 0,linetype="dashed")+
  theme_bw()+
  theme(legend.title = element_blank(),
        legend.position = "none",
        aspect.ratio = 1,
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(breaks = c("Reallocation: No -  Likelihood: Yi",
                               "Reallocation: Yes -  Likelihood: Yi*",
                               "Reallocation: Yes -  Likelihood: Dj"),
                    values=c("gold","#F8766D","#00BA38"))+
  xlab("")+ylab(parse(text = TeX('$\\mu$')))+
  ggtitle("Intercept of the latent field")+
  geom_hline(yintercept = 2,col="red",linetype="dashed")

@@

@@

The $D_j$ model allows to recover the species-habitat relationship and to improve the accuracy of the spatial predictions (Figure \@ref(fig:PerfMetricSingle)). Still, the $D_j$ model outputs are not as accurate as the ones of the golden-standard. Furthermore, the zero-inflation parameter is unbiased when the model is fitted to catch declarations $D_j$. Other parameters (observation variance, intercept) are also better estimated even though they remain slightly biased (Figure \@ref(fig:ParBiasSingle)). This alternative model have some convergence difficulties (Table 2) as 8% of the model did not converged when sample size is medium (100 pings) and only 3% did not when sample size is large (1000 pings).

![(\#fig:PerfMetricSingle) Performance metric for single-square simulations. Columns: number of commercial fishing positions. x-axis: number of zones visited within each declaration. 'Reallocation: ', data are or are not reallocated in simulations. 'Likelihood: ', the likelihood is computed on exact punctual observations $Y_i$, on reallocated observations $Y_i^*$ or on catch declarations $D_j$. Gold: golden standard. Red: uniform reallocation ($Y_i^*$ model). Green: model-based reallocation ($D_j$ model). Simulations conducted on 10 fishing positions are not represented as they mostly did not converged for the $D_j$ model (see Table 2).](images/Perf.metric_single_square.png){width=75%}

![(\#fig:ParBiasSingle) Parameters relative bias for single-square simulations. 'Reallocation: ', data are or are not reallocated in simulations. 'Likelihood: ', the likelihood is computed on exact punctual observations $Y_i$, on reallocated observations $Y_i^*$ or on catch declarations $D_j$. Gold: golden standard. Red: uniform reallocation ($Y_i^*$ model). Green: model-based reallocation ($D_j$ model). Only the simulations with 1000 fishing positionss are represented. Black line: zero value. Red line: parameter true value.](images/par_plot_single_square.png)

@@

knitr::kable(Results_conv_2,booktabs = T,align = "c",
                  caption = "Single-square simulations - Percentage of convergence per simulation-estimation configuration.")

## Multiple-square simulations

@@

Results_conv_2$aggreg_obs[which(Results_conv_2$aggreg_obs == "Yi")] <- "Yi*"

@@

Results_conv_2 <- Results_conv_2[-which(Results_conv_2$Model == "Scientific model" &
                                          Results_conv_2$`Likelihood level` == "Dj"),]
Results_conv_2$`Likelihood level`[which(Results_conv_2$Model == "Scientific model" &
                       Results_conv_2$`Likelihood level` == "Yi*")] <- " "

@@

Results_plot$obs <- NA
Results_plot$obs[which(Results_plot$aggreg_obs == T)] <- "Dj"
Results_plot$obs[which(Results_plot$aggreg_obs == F)] <- "Yi*"
Results_plot$obs[which(Results_plot$Estimation_model == 2)] <- "Scientific data"
Results_plot$obs <- factor(Results_plot$obs,levels = c("Yi*","Dj","Scientific data"))

@@

beta_plot <- ggplot()+
  geom_boxplot(data = Results_plot,
               aes(x = Model,
                   y = beta1_est,
                   fill = obs))+
  geom_hline(yintercept = 0,linetype="dashed")+
  theme_bw()+
  theme(legend.title = element_blank(),
        legend.position = "none",
        aspect.ratio = 1)+
  # facet_wrap(.~simu_type)+
  ylab(TeX("$\\beta_S$")) +
  xlab("")+
  geom_hline(yintercept = 2,col="red",linetype="dashed")+
  scale_color_manual(breaks = c("Scientific model",
                                "Yi*",
                                "Dj"),
                     values=c("#619CFF","#F8766D","#00BA38"))
